Univariate Analysis

### 2017-08-22 model fitting/univariate_analysis.Rmd
reformatted the file

# Univariate analyses (select variables with p-value < 0.01)
#===================================================================
# category 
pvalues <- get_pvalue("category")
#===================================================================
# new_category
pvalues <- get_pvalue("new_category")
#===================================================================
# n_images
pvalues <- get_pvalue("n_images")

@@

#===================================================================
# n_tags
pvalues <- get_pvalue("n_tags", factor = TRUE)

#===================================================================
# new_user
pvalues <- get_pvalue("new_user")

#===================================================================
# weekday
pvalues <- get_pvalue("weekday")

#===================================================================
# hour of the day
pvalues <- get_pvalue("hour", factor = TRUE)

#===================================================================
# morning/noon/evening/night?
pvalues <- get_pvalue("ampm")

#===================================================================
# title length
pvalues <- get_pvalue("title_length")

#===================================================================
# text length
pvalues <- get_pvalue("text_length")

#===================================================================
# device name length
pvalues <- get_pvalue("device_length")

#===================================================================
# length of sentence until first punctuation mark
pvalues <- get_pvalue("text_till_punct")

#===================================================================
# if text contains any end punctuation
pvalues <- get_pvalue("text_contain_punct")

#===================================================================
# if the title ends with a question mark
pvalues <- get_pvalue("title_questionmark")

#===================================================================
# if title begins with "wh"
pvalues <- get_pvalue("title_beginwh")

#===================================================================
# if first letter of title is capitalized
train1$capital_title <- str_detect(as.character(train1$title), pattern = "^[[:upper:]]")
pvalues <- get_pvalue("capital_title")

#===================================================================
# if first letter of text is capitalized
train1$capital_text <- str_detect(as.character(train1$text), pattern = "^[[:upper:]]")
pvalues <- get_pvalue("capital_text")

#===================================================================
# if text is in all lower case
pvalues <- get_pvalue("text_all_lower")

#===================================================================
# prior_effort
pvalues <- get_pvalue("prior_effort")

#===================================================================
# expressing gratitude 
pvalues <- get_pvalue("gratitude")

#===================================================================
# greeting
pvalues <- get_pvalue("greeting")

#===================================================================
# update the question?
pvalues <- get_pvalue("update")

#===================================================================
# newline ratio to length of text
pvalues <- get_pvalue("newline_ratio")
```

```{r, tag based and frequent term variables, include = FALSE}
# tag-based variables 
#===================================================================
# avg_tag_length
pvalues <- get_pvalue("avg_tag_length")

#===================================================================
# avg_tag_score
pvalues <- get_pvalue("avg_tag_score")

#===================================================================
# number of "frequent" tags a question contains
pvalues <- get_pvalue("num_freq_tags")

#==================================================================================
#frequent terms in unanswered/answered questions
pvalues <- get_pvalue("contain_unanswered")
pvalues <- get_pvalue("contain_answered")
```

### researching splines + quadratic terms + square root transformations

* spline: piecewise m - 1 degree polynomial that is continuous up to its first m - 2 derivatives
* obtain more flexible curves: increase the degree of spline/add more knots
* tradeoff: 
    + few knots/low degree: class of functions may be too restrictive (bias)
    + many knots/high degree: risk overfitting 
* spline has m + K degrees of freedom 
* types of splines
    + cubic splines (most common)
* truncated power basis - flawed
* B-spline basis: much more stable, typically used 
* Natural cubic splines
    + polynomial fits tend to be erratic at boundaries of data (even worse for cubic splines)
    + natural cubic splines fix this by adding 4 constraints that the function is linear beyond the boundaries of the data 
    + because basis functions are fixed, model fitting and inference are straightforward exentsions to how the model would normally be without splines 
* **use pspline and penalized spline smoothing method with AIC criteria**

```{r, echo = FALSE}
# fit splines + sqrt/log transformations on title_length, text_length, device_length, avg_tag_length, newline_ratio

# new function to add p-values from these fits to pvalue data frame
get_pvalue_new <- function(variable, fit) {
  summary <- summary(fit)
  df <- data.frame(variable = variable, pvalue = unname(summary$logtest[[3]]))
  pvalues <- rbind(pvalues, df)
  return(pvalues)
}

>>>>

# Univariate analyses (select variables with p-value < 0.01)
#===================================================================
pvalues <- get_pvalue("category")
pvalues <- get_pvalue("new_category")
pvalues <- get_pvalue("n_images")

@@

pvalues <- get_pvalue("n_tags", factor = TRUE)
pvalues <- get_pvalue("new_user")
pvalues <- get_pvalue("weekday")
pvalues <- get_pvalue("hour", factor = TRUE)
pvalues <- get_pvalue("ampm")
pvalues <- get_pvalue("title_length")
pvalues <- get_pvalue("text_length")
pvalues <- get_pvalue("device_length")
pvalues <- get_pvalue("text_till_punct")
pvalues <- get_pvalue("text_contain_punct")
pvalues <- get_pvalue("title_questionmark")
pvalues <- get_pvalue("title_beginwh")

# if first letter of title is capitalized 
train1$capital_title <- str_detect(as.character(train1$title), pattern = "^[[:upper:]]")
pvalues <- get_pvalue("capital_title")

# if first letter of text is capitalized
train1$capital_text <- str_detect(as.character(train1$text), pattern = "^[[:upper:]]")
pvalues <- get_pvalue("capital_text")

pvalues <- get_pvalue("text_all_lower")
pvalues <- get_pvalue("prior_effort")
pvalues <- get_pvalue("gratitude")
pvalues <- get_pvalue("greeting")
pvalues <- get_pvalue("update")
pvalues <- get_pvalue("newline_ratio")
pvalues <- get_pvalue("avg_tag_length")
pvalues <- get_pvalue("avg_tag_score")
pvalues <- get_pvalue("num_freq_tags")
pvalues <- get_pvalue("contain_unanswered")
pvalues <- get_pvalue("contain_answered")
```

### splines + quadratic/cubic terms + square root/log transformations

* **use pspline and penalized spline smoothing method with AIC criteria**

```{r, echo = FALSE}
# new function to add p-values from these fits to pvalue data frame
get_pvalue_new <- function(variable, fit) {
  summary <- summary(fit)
  df <- data.frame(variable = variable, pvalue = unname(summary$logtest[[3]]))
  pvalues <- rbind(pvalues, df)
  return(pvalues)
}
